Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses

### 2014-12-06 doc/slides/hong+eisner.TLT13.slides.tex
initial outline of paper

%
% File acl2014.tex
%
% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
%%
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[a4paper,11pt]{article}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}         
\usepackage{pslatex}

\usepackage{hyperref}
\pagestyle{empty}

%\bibliographystyle{acl2014}
\bibliographystyle{plain}
\newcommand{\newcite}[1]{\cite{#1}}   % ACL's newcite is not available in plain bib style, so just use cite

%\documentclass[11pt]{article}
%\usepackage{acl2014}
\usepackage{times}
%\usepackage{url}
%\usepackage{latexsym}
\usepackage{amsmath}

%\usepackage{xspace}
\usepackage{enumitem}
\usepackage[small]{caption}
\usepackage{longtable}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
%\usepackage{soul}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
%\usepackage{caption}
\usepackage{subcaption}

@@

\newcommand{\eqnref}[1]{\eqref{eqn:#1}}

\usepackage[disable]{todonotes}   % insert [disable] to disable all notes
% Note that these macros accept optional arguments such as 
% [size=\small,bordercolor=red].
\newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
\newcommand{\noteJH}[2][]{\Note[size=\small,#1]{JH}{blue!40}{#2}}
\newcommand{\noteJE}[2][]{\Note[size=\small,#1]{JE}{green!40}{#2}}   
\newcommand{\notewho}[3][]{\Note[size=\small,#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
\newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
\newcommand{\NoteJE}[2][]{\noteJE[inline,#1]{#2}}
\newcommand{\Notewho}[3][]{\notewho[inline,#1]{#2}{#3}}  % extra arg with miscellaneous author

\title{Deriving Multi-Headed Planar Dependency Parses from Link
  Grammar Parses\thanks{This material is based upon work supported by the National
    Science Foundation under Grant No. 1423276.  The work was mainly
    conducted while the first author was at Johns Hopkins
    University.\noteJE{anyone else to thank?}}}
% Oriented link grammar: Creating a multi-headed dependency corpus.

\author{
  Juneki Hong \\
  [0.5cm] Language Technologies Institute\\
  Carnegie Mellon University\\
  Pittsburgh, PA 15213, USA\\
  {\texttt{juneki@cs.cmu.edu}}
  \and
  Jason Eisner\\
  [0.5cm] Department of Computer Science\\
  Johns Hopkins University\\
  Baltimore, MD 21218, USA \\
  {\texttt{jason@cs.jhu.edu}}
}

\date{}

%\setlength\titlebox{6.5cm}    % Expanding the titlebox

\begin{document}
\maketitle

\begin{abstract}
\noindent
Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective or planar case.
% OLD
% investigate the benefit of such parsers or to work on making them faster or more accurate.  
To help fill this gap, we observe that link grammar already produces {\em undirected} planar graphs, and so we wanted to determine whether these could be converted into directionalized dependency parses.
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.  We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. 
The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.

\end{abstract}

\thispagestyle{empty}

%\newpage

\section{Motivation}\label{sec:motivation}

Link Grammar \cite{sleator-temperley-1991} is a syntactic formalism in which a parse of a sentence is an undirected, edge-labeled, planar graph.  The labeled edges of the graph represent syntactic relationships among the words.  The vertices of the graph are simply the words $1, 2, \ldots, n$ of the sentence, along with a distinguished ``root'' vertex 0.

The small Link Grammar community has invested effort in creating link grammars for several languages.  In this short paper, we consider whether their {\em undirected} parses can be converted automatically to {\em directed} ones.  We have three motivations:

\begin{enumerate}

\item We were curious about the relation between link grammar annotation and dependency grammar annotation.  We suspected that the link grammar parses could be interpreted as multi-headed dependency grammar parses.  Although the link grammar authors did not bother to specify directions for the different edge types, we suspected that they essentially had such directions in mind.

\item Our problem provides a good case study for how to automatically enrich a corpus.  Hand-constructed grammars or corpora sometimes provide a lighter level of annotation than desired.  In our setting, the edges lack directions; in other settings, the syntactic categories may be coarser than desired, or some relations may be omitted.  One may want to automatically enrich the annotation in such cases, whether by doing some kind of learning \cite[et seq.]{matsuzaki-et-al-2005}, or by exploiting implicit constraints.  In this paper, we use Integer Linear Programming to exploit implicit constraints of consistency and acyclicity.

\item The resulting parses may be useful data for experimenting with new parsing {\em algorithms}.  There has been a good deal of recent research on projective dependency parsing, variously using global optimization or sequential classification (see \cite{martins-et-al-2008,bohnet-2011,chen-li-zhang-2014} for surveys). Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic and computational interest for reasons discussed below.  However, for training and testing such algorithms, one would need a plausible sample of multi-headed projective dependency parses of real sentences.  Our method cheaply manufactures such a sample, to compensate for the current lack of gold-standard data of this form.

\end{enumerate}

\noindent Our automatic method for reconstructing the latent directions also had an unexpected benefit.  It revealed an inconsistency in the hand-written English link grammar, regarding the handling of embedded sentences with missing (PRO) subjects.

\section{Multi-Headed Dependency Parsing}

Dependency parsing maps a sentence to a directed graph whose vertices are the words $1, 2, \ldots, n$ of the sentence along with a distinguished ``root'' vertex 0.  A labeled directed edge $u \stackrel{L}{\rightarrow} v$ or $v \stackrel{L}{\leftarrow} u$ indicates that the ``child'' $v$ is some kind of argument or modifier of its ``parent'' $u$.  The edge label $L$ indicates the specific syntactic or semantic relationship between the two words.

In the special case $u=0$, the edge designates $v$ as playing some special top-level role in the sentence, e.g., as the main verb.  We disallow $v=0$.

As discussed by \newcite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
\begin{itemize}[noitemsep]
\item {\sc Single-Head}: Each word has $\leq 1$ parent.
\item {\sc Acyclic}: There are no directed cycles.
\item {\sc Connected}: Each pair of words has a undirected path between them.
\item {\sc Reachable}: Each word can be reached from 0 by a directed path (which implies {\sc Connected}).  Note that 0 may have multiple children.
\item {\sc Planar}: edges may not ``cross.''  That is, if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then {\sc Planar} requires $i \leq v \leq j$.
\end{itemize}
It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.

In this paper, we are interested in relaxing the {\sc Single-Head} requirement while preserving all the others.  This is the setting of {\em multi-headed projective dependency parsing}.  Just as in the single-headed case, the other requirements ensure that all edges are projective.  (A projective edge is one where the parent is an ancestor of all words between the parent and the child \cite{nivre-slides}.)

Relaxing {\sc Single-Head} means that the parse can have more than $n$ edges, allowing it to express more relationships between words.  
In English, for example, here are some constructions that seem to call for a multi-headed analysis:
\begin{description}
\item[control] In {\em ``Jill likes to skip,''} the word {\em Jill} is the subject of two verbs.  In {\em ``Jill persuaded Jack to skip,''} {\em Jack} is the object of one verb and the subject of another.  Without recognizing this, our parser would miss the syntactic invariants that {\em skip} always has a subject and {\em persuaded} always has an object.  It would also be unable to exploit the selectional preferences of both verbs to help disambiguate the parse.  This is why we prefer to make the parser aware of multi-headedness, rather than using a single-headed parser and then extracting the additional semantic roles from its output.
\item[relativization] In {\em ``The boy that Jill skipped with fell down,''} the word {\em boy} is the object of {\em with} as well as the subject of {\em fell}.  Without recognizing this, we would miss the syntactic invariant that {\em with} always has an object.  
\item[conjunction] In {\em ``Jack and Jill went up the hill,''} {\em Jack} and {\em Jill} serve as the two arguments to {\em and}, but they are also semantically subjects of {\em went}.  Without recognizing this, we would have no (local) reason for expecting the arguments of {\em and} to be nouns.
\end{description}

\noindent In linguistics, it is common to analyze some of these structures using trees with ``empty categories.''  The subject of {\em skip} is taken to be a silent morpheme {\em PRO}:
{\em ``Jill$_i$ likes PRO$_i$ to skip.''}  However, this is no longer a tree if one considers the implicit undirected edge between {\em Jill} and {\em PRO} (denoted by their shared index $i$).  Our simpler representation contracts this coreference edge, eliminating {\em PRO} and creating a {\em Jill} $\leftarrow$ {\em skip} link.

\medskip

An anonymous reviewer objects to research on projective parsing algorithms, since the {\sc Planar} restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.  However, efficient algorithms often exist in the projective case, and these projective algorithms---which are typically first developed on a projective corpus of the sort that we will construct---can be useful even when the true parses are not quite projective.
Natural-language parses have a low rate of non-projective edges \cite{nivre-slides} and the non-projective parses tend to be ``almost projective'' \cite{pitler-et-al-2013}.  Thus, one can apply a fast projective parser as an approximate method, or as one ingredient in a more complex model \cite{martins-et-al-2013}, or as the first step in a coarse-to-fine or stacking architecture \cite{rush-petrov-2012,martins-et-al-2008} in order to obtain preliminary edge scores that are then supplied to a non-projective parser.  Another approach is to transform non-projective parses into a projective annotation style so that projective parsers can be used \cite{nivre-nilsson-2005,nivre-2009}.

\section{Link Grammars}

Graph representations of syntactic and semantic structure have been widely considered of late \cite{demarneffe-manning-2008,demarneffe-et-al-2013,ivanova-et-al-2012,AMR-2013,oepen-et-al-2014}.  A few past NLP papers have explored multi-headed dependency parsing \cite{buchkromann-2006,mcdonald-pereira-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}.  They constructed their multi-headed dependency corpora by automatically converting from other formats such as HPSG.  Currently there seem to be no corpora that were directly annotated in this form, other than the Danish Dependency Treebank \cite{kromann-2003}.

% ADDRESSED BY ABOVE?
% \NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).  However, that requires per-language}

The above work considered non-projective parses.  It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.  However, this is only half-true.  Link Grammar \cite{sleator-temperley-1991} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.  It produces undirected connected planar graphs.  Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser \cite{LINK-1999}.  The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html},
which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.  As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.  For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.  Each entry has a hand-assigned cost in \{0,1,2\}, and the parser finds the parse of minimum total cost \cite{LINK-1999,vepstas-2014}.

Given a link grammar parse, it would be straightforward to convert it to an acyclic dependency parse by orienting all edges rightward.  However, the result may violate the {\sc Reachable} constraint.  Instead we could orient all edges by depth-first search from the root node, which yields a DAG satisfying all our constraints.  However, this might result in inconsistent annotation conventions, with some \texttt{S}-labeled links pointing from subject to verb and others from verb to subject.

In the English link grammar, an \texttt{S} edge encodes a ``subject-verb'' relation {\em whose left word serves as the subject}.  We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links should be interpreted as pointing leftward (from verb to subject: {\em ``\mbox{Jack $\stackrel{\texttt{S}}{\leftarrow}$ is} falling''}).

In general, we supposed that the link grammar lexicon designers actually had a {\em consistent} direction in mind for each {\em edge label}.  This does not imply that English subjects must always appear to the left of the verb!  The link grammar designers took care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``\mbox{Is $\stackrel{\texttt{SI}}{\rightarrow}$ Jack} falling?''}).  Similarly, different edge labels are used for English ``object-verb'' relations according to whether it is the left or the right word that serves as the object.  These labels are presumably intended to encode different edge directions.

% PREVIOUS ATTEMPT
% This does not rule out variation in the word order used to express a semantic relation.  For example, the English link grammar allows both SVO and VSO sentences, but with different edge labels.  An \texttt{S} edge encodes a ``subject-verb'' relation whose the left word serves as the subject, while an \texttt{SI} edge encodes the same relation when the the right word serves as the subject.

Our goal in this paper is to recover these implicit directions by global optimization.  We seek a fixed mapping from labels to directions such that link grammar parses become directed dependency parses that satisfy all of our constraints.

Our first thought was to seek a direction mapping such that no parsed word sequence allowed by the link grammar lexicon could possibly violate our constraints after directions were imposed.  This is a well-defined constraint programming problem.  For example, to prevent cyclicity, we would require (roughly speaking) that no word type in the lexicon could follow a sequence of directed rightward links through other word types and then a leftward link back to itself.

% REMOVED AT REVIEWER SUGGESTION
% However, working directly with the link grammar lexicon format is somewhat tricky. We also feared
However, we feared that there would not be a feasible solution---because of errors in the lexicon or linguistically unnatural word sequences not anticipated by the grammar designers.  In this case it would be unclear how to relax our constraints.

Thus, rather than considering all theoretically possible word sequences, we chose to use a sample of {\em naturally occurring} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed. 
If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.  In this way, the empirical distribution of sentences guides our assignment of directions.  We are releasing the resulting multi-headed directed corpus via our personal websites. \noteJE{is this the best way?}

% Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word. A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity. The resulting links describe the relationships between constituents in a parse.

% The link grammar is based on a set of handwritten dictionaries. Instead of going through these dictionaries, we learned the grammar using an ILP. This approach also potentially allows us to analyze link grammar dictionaries other than English. \todo{explore other link grammar dictionaries}

% \NoteJE{what statistics does it use?  what languages are available}

% Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies, and whether the multi-headedness of the link grammar offers additional dependency relationships not found in other corpora.

% Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree. Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies. To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \todo{}.

\section{Data Sets}
We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}---a subset of the Penn Treebank for which {\em single}-headed reference parses are available. 
%We also used a prefix of the Russian National Corpus,\footnote{That is, a prefix when considering the files in Unix sort order by filename.  We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.} which is unparsed.
We also used a prefix of the Russian News Commentary data from the ACL 2013 Shared Task of Machine Translation,\footnote{\url{http://www.statmt.org/wmt13/training-monolingual-nc-v8.tgz}} which is unparsed.

We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.  The parser's coverage is less than ideal: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \mbox{\input{figure/russian_original_sentence_count.tex}$\!\!$)} Russian sentences, discarding the other sentences.\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a time limit. We removed these sentences for fear that the parses would be unreliable.
% REMARK BY JUNEKI, PROBABLY TOO DISTRACTING
% our ILP encoding tries to satisfy the edges of an entire connected parse. Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus. It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.
}
These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.

On English, the link grammar parses have \input{figure/conll_analysis_edge_percent.tex}more edges overall, indicating that their directed versions will have a few multi-headed constructions per sentence.  They
do differ in style from the single-headed CoNLL parses of the same English sentences.  Only \input{figure/conll_analysis_precision.tex}of the links match CoNLL arcs, and only \input{figure/conll_analysis_recall.tex}of the CoNLL arcs match links.
% DATA FOR THE ABOVE
% \begin{figure}[ht!]
%   \centering
%   \small
%   \input{figure/conll_analysis_arcs.tex}
%   \caption{Arcs that match links}
% \end{figure}
% 
% OLD
% Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus. \todo{This is because...}

% With the English sentences we compared our link parse results to the original dependency annotations, and with the Russian sentences we only produced link parses.

%\subsection{Discarding Incomplete Link Parses}

%\NoteJE{ideally we could cut this section given what I wrote above.  However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.  Didn't we decide to drop those sentences too?}

%Not all of the data was used for this work. From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root. \input{figure/conll_analysis_sentences_disconnected} of the English sentences were removed in this way. The ILP was run only on the connected parses to produce a corpus of directionalized link parses.

%For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence. About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences were removed, leaving us working only with complete link parses.

%\begin{figure}[ht!]
%  \centering
%  \small
%  \input{figure/conll_analysis_sentences.tex}
%  \caption{\small The English sentences used for this work.}
%\end{figure}

%\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}

\section{Integer Linear Programming Model}

For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens in the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.  We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}

% OLD VERSION
% we introduce a variable $x_{ij}$ that is 0 or 1 according to whether $e$ points left or right.  We abbreviate $1-x_{ij}$ as $\bar{x}_{ij}$.

For each non-0 token $v$, we ensure that it has at least one parent by constraining\footnote{To denote two linked tokens, we use variables $i,j$ when $i$ is to the left of $j$, or variables $u,v$ when $u$ is the parent of $v$.}
\begin{align}\label{eqn:oneparent}
\sum_u x_{uv} & \geq 1
\end{align}
% OLD VERSION
% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
where $u$ ranges only over tokens such that the relevant variable exists.
%
To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.  We require a child's depth to be at least 1 greater than each of its parents' depths---constraints that can be satisfied iff the sentence has no directed cycles:
\begin{align}\label{eqn:nocycles}
(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
% OLD VERSION
% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
\end{align}
The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  
\NoteJE{in final version: figure out whether it is faster to require the 0 tokens to have depth 0, or add all the depths (perhaps multiplied by 0.001) to the objective function to break ties in simplex, or neither.  See also other notes in tex file.}
% As a speedup, we can require the 0 tokens to have depth 0. \noteJE{does this really still make a difference to speed?  
% \noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}

Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.  For each label $L$, we introduce binary variables $r_L$ and $\ell_L$, which say whether a link of type $L$ is ``allowed'' to point right or left, respectively.  For each undirected edge $ij$ of label $L$, with $i < j$, we write
\begin{align}
x_{ij} &\leq r_L + s_{ij} & 
x_{ji} &\leq \ell_L + s_{ij}
\end{align}
where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.

Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
\begin{align}\label{eqn:obj}
\min \left( \sum_{L} r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij} s_{ij}
\end{align}
where $N_L$ is the number of link tokens with label $L$.  Objective \eqnref{obj} is willing to tolerate up to 1/4 of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions. 
\noteJE{check!}
One could experiment with adjusting the constant 1/4; we selected that value simply because it seemed like a reasonable threshold {\em a priori}.
% CUT FOR SPACE
% (i.e., $r_L+\ell_L=2$).

\NoteJE{It might be a good idea to have the second allowed direction still carry some lesser slack penalty, so that we break ties in favor of the preferred direction.  The two bits for each label would then say ``which direction is preferred?'' and ``is the other direction allowed?''}

% OLD
% We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root. Our task of finding the smallest set of assignments that would explain the parses is also called model minimization\cite{ravi2009}. \noteJE{worth citing, but not quite the same thing}
% For every encountered label we generate two boolean label/direction variables (e.g. label/left, label/right). Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
% The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints. The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
% \subsection{Slack}
% We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}. This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.

% 
% \begin{figure*}[ht!]
%   \vspace{-6pt}
%   \centering
%   \input{figure/sample_parses.tikz}
%   \vspace{-6pt}
% \end{figure*}

\section{Experiments and Results}

%\hspace*{-1.5in}
\begin{figure*}[t]

  \vspace{-6pt}
  \centering

  %\input{figure/sentences.tikz}

>>>>

\documentclass{beamer}

% For more theme options:
%/usr/share/texmf/tex/latex/beamer/base/themes/theme/compatibility
\usepackage{beamerthemelined}

% this package seems to throw an error for me. -Juneki 12/6/14
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{soul}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

\newcommand{\eqnref}[1]{\eqref{eqn:#1}}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
\usepackage{todonotes}   % insert [disable] to disable all notes
\newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
\newcommand{\noteJH}[2][]{\Note[#1]{JH}{blue!40}{#2}}
\newcommand{\noteJE}[2][]{\Note[#1]{JE}{green!40}{#2}}
\newcommand{\notewho}[3][]{\Note[#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
\newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
\newcommand{\NoteJE}[2][]{\noteJE[inline,#1]{#2}}
\newcommand{\Notewho}[3][]{\notewho[inline,#1]{#2}{#3}}  % extra arg with miscellaneous author

\begin{document}
\title{Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses}
\author{Juneki Hong, Jason Eisner}
\date{\today}
\frame{\titlepage}

\section{Motivation}
\frame{\frametitle{Motivation}

}

\section{Multi-headed}
\frame{\frametitle{Multi-headed Dependency Parsing}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{multiheaded}
  \caption{English is not completely multiheaded}
  \NoteJH{image taken from Chris Dyer's slide: http://demo.clab.cs.cmu.edu/fa2014-11711/images/0/0e/Depparsing.pdf}
\end{figure}

}

\section{Link Grammars}
\frame{\frametitle{Link Grammars}

\begin{figure}
\begin{subfigure}[b]{0.4850\textwidth}
	\begin{dependency}
		\begin{deptext}
			the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
			- \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
		\end{deptext}
		\depedge[edge above, edge style = {red}]{2}{3}{S}
		\depedge[edge above, edge style = {red}]{3}{6}{I}
		\depedge[edge above, edge style = {red}]{2}{1}{D}
		\depedge[edge above, edge style = {red}]{7}{8}{MV}
		\depedge[edge above, edge style = {red}]{6}{7}{P}
		\depedge[edge above, edge style = {red}]{8}{9}{J}
		\deproot[edge above, edge style = {red}]{2}{W}
		\deproot[edge above, edge style = {red}]{7}{WV}
		\deproot[edge above, edge style = {red}]{10}{X}
		\depedge[edge above, edge style = {red}]{6}{4}{E}
		\depedge[edge above, edge style = {red}]{6}{5}{E}
	\end{dependency}
\end{subfigure}

\end{figure}
\NoteJH{figure not centered}
\NoteJH{Make the edges undirected}

}

\frame{\frametitle{Link Grammars}
\begin{figure}
\begin{subfigure}[b]{0.4850\textwidth}
	\begin{dependency}
		\begin{deptext}
			{\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
			the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
			- \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
		\end{deptext}
		\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
		\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
		\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
		\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
		\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
		\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
		\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
		\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
		\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
		\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
		\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
		\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
		\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
		\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
		\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
		\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
		\deproot[edge below, edge style = {red, dotted}]{2}{W}
		\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
		\deproot[edge below, edge style = {red, dotted}]{10}{X}
		\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
		\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
	\end{dependency}
\end{subfigure}

\end{figure}

}

\section{ILP}
\frame{\frametitle{Integer Linear Programming}

\NoteJH{Will the audience know about ILP?}

}

\section{ILP Model}
\frame{\frametitle{Integer Linear Programming Model}

\NoteJH{Take from the paper.}

}

\section{Experiments and Results}
\frame{\frametitle{Experiments and Results}

}

\section{Conclusions}
\frame{\frametitle{Conclusions}

}
